Supplement 1: Different Response Distributions

### 2018-09-29 pkg/disttree/inst/draft/supplement/supplement_distributions.Rnw
more work on supplements
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2299 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 42375fce433cf91b5ec4e546daaa344ee5df0bc4

\item \emph{Hurdle model \citep[or two-part model,][]{Cragg:1971,Mullahy:1986}:}
An additional parameter $\nu$ describing the probability that there is 
any precipitation at all, i.e. $\nu = \mathbb{P}(Y>0)$, is included in the model.
The three-parameter log-likelihood function for the full hurdle model 
with parameter vector $\bm{\theta} = (\mu, \sigma, \nu)$ is
\begin{equation} \label{eq:hurdle}
\ell_{\mathrm{hurdle}}(\mu, \sigma, \nu; Y) = 
\begin{cases}
\log\left\{\nu \cdot \frac{f(Y; \mu, \sigma)}{1 - F(0; \mu, \sigma)} \right\}, & \text{if } Y > 0 \\[0.2cm]
\log(1 - \nu), & \text{if } Y = 0.
\end{cases}
\end{equation}
For $Y=0$ the probability for no rain is expressed by $1-\nu$. 
For $Y>0$ the probability that it will rain $\nu$ is multiplied with the 
left-truncated at zero density function. In contrast to the censored model, 
there is thus a dedicated parameter $\nu$ controlling the point mass while 
the other parameters $\mu$ and $\sigma$ only influence the distribution of
the positive observations.

In the paper a left-censored at zero Gaussian distribution is chosen.
Here, we additionally apply a left-censored at zero logistic distribution
and a hurdle model based on a Gaussian distribution.

In the log-likelihood function in Equation~\ref{eq:hurdle} properties of the
logarithmic function can be applied to split the term for $Y>0$ 
into two additive terms separating the parameters $\mu$ and $\sigma$
from the additional parameter $\nu$. Therefore, $\nu$ can be modeled 
by a separate model in the GAMLSS framework. The same approach is
chosen for the distributional forest model in this application study.

The models are compared based on the continuous ranked probability 
score (CRPS). For the censored models the closed form following 
\cite{Jordan:2018} is used while a numerical approximation is 
calculated for the hurdle models.

\section{Stations}
Out of the 95 observation stations 15 are considered for this benchmark 
study (see Figure~\ref{fig:map}). 
They have been selected based on the region they are in, their altitude 
and the performance of the different models in the application in the 
paper such that all of these features are represented equally.

Selected stations: Axams, Lech, Zuers, See im Paznaun, Jungholz, Ladis-Neuegg, 
Oetz, Ochsengarten-Obergut, Ginzling, Rotholz, Walchsee, Koessen, 
Innervillgraten, Matrei in Osttirol, St.Johann im Walde.

\begin{figure}[h]
\vspace*{-3cm}
\setkeys{Gin}{width=0.99\textwidth}
<<map_stationwise, fig=TRUE, echo=FALSE>>=
##  map
data("StationsTyrol", package = "RainTyrol")
data("MapTyrol", package = "RainTyrol")
library("sp")
sp <- SpatialPointsDataFrame(subset(StationsTyrol, select = c(lon, lat)),
                             data = subset(StationsTyrol, select = -c(lon, lat)),
                             proj4string = raster::crs(MapTyrol$RasterLayer))

plot(MapTyrol$SpatialPolygons)
points(sp, pch = 21, col = "darkgrey", las = 1, cex = 0.6)
points(sp[c(5, 6, 20, 23, 25, 32, 33, 46, 47, 56, 57, 70, 79, 82, 83),], 
       pch = 21, bg = hcl(325, 100, 70), cex = 1.5)
# stationname beneath
text(sp[c(25, 47, 57, 70),], 
     labels = StationsTyrol$name[c(25, 47, 57, 70)],
     pos=1, cex=0.75)
# stationname left
text(sp[c(20, 32, 46, 79, 83),], 
     labels = StationsTyrol$name[c(20, 32, 46, 79, 83)],
     pos=2, cex=0.75)
# stationname above
text(sp[c(6, 33, 82),], 
     labels = StationsTyrol$name[c(6, 33, 82)],
     pos=3, cex=0.75)
# stationname right
text(sp[c(5, 23, 56),], 
     labels = StationsTyrol$name[c(5, 23, 56)],
     pos=4, cex=0.75)
@
\vspace*{-3.5cm}
\caption{\label{fig:map}Map of Tyrol with all 95 observation stations and 
the 15 considered stations highlighted in pink.}
\end{figure}

\section{Results}
In Figure~\ref{fig:lattice} each panel belongs to one of the 15 observation stations. 
The three distributions \textit{gaussian} (left-censored at zero Gaussian distribution), 
\textit{hgaussian} (hurdle model based on the Gaussian distribution) and 
\textit{logistic} (left-censored at zero logistic distribution) are compared based 
on the resulting CRPS values for each of the four applied methods that are represented 
by different colors and symbols. 
Overall, the forest model performs almost equally well for all three distributions. 
Therefore it seems to be very robust against missspecifications of the distribution.
For the two GAMLSS models the censored Gaussian and the hurdle model perform similarly 
at most of the stations while applying the censored logistic distribution leads to an 
increase in CRPS, i.e. to a decline in performance, especially for the prespecified 
GAMLSS (for example at stations Zuers and Jungholz).
For the EMOS model exchanging the distribution does not have a strong effect on its 
performance, similar to the forest model.
All in all, exchanging the underlying distribution does not lead to clear 
improvements of any of the methods, especially the censored Gaussian and the 
hurdle model achieve almost the same results. However, the robustness against 
missspecification of the distribution varies slightly over stations. For example 
at station Matrei in Osttirol all methods seem to be very stable while stronger 
influences can be seen at station Jungholz.

\textit{Note:} When applying a hurdle model for the stations Oetz and 
Innvervillgraten the prespecified GAMLSS computationally failed to build/fit 
a model. Therefore this method is not represented in the corresponding panels 
of Figure~\ref{fig:lattice} for these two cases.

In Figure~\ref{fig:box_crpss_gaussian} the tested distributions are compared 
for each method based on CRPS skill score with the censored Gaussian distribution 
as reference represented by the horizontal line at zero. Again, the forest model 
and the EMOS show only slight variations (on average not even one percent) over different 
distributions while the two GAMLSS models perform even or slightly better applying 
a hurdle model and loose strength when employing a logistic distribution compared 
to a Gaussian distribution.

\begin{figure}
\setkeys{Gin}{width=0.99\textwidth}
<<results_lattice, fig=TRUE, echo=FALSE, results=hide, height=10, width=8>>=
library("lattice")
##### 
# HCL palette
pal <- hcl(c(10, 128, 260, 290, 50), 100, 50)

load("Rain_distributions.rda")
stations <- c("Axams", "Lech", "Zuers", "See im Paznaun", "Jungholz", 
              "Ladis-Neuegg", "Oetz", "Ochsengarten-Obergut",
              "Ginzling", "Rotholz", "Walchsee", "Koessen", 
              "Innervillgraten", "Matrei in Osttirol", 
              "St.Johann im Walde")

@@

xyplot(crps ~ distribution | station, groups = ~ method, 
       data = means, auto.key = TRUE, 
       type = "o", lwd = 2, lty = 1,  
       drop.unused.levels = TRUE, layout = c(3,5),
       as.table = TRUE, ylab = "CRPS", xlab = "Distribution")
@
\caption{\label{fig:lattice}CRPS values of all four methods, once applied with 
each of the three distributions for each of the 15 observation stations.}
\end{figure}

\begin{figure}
\setkeys{Gin}{width=0.99\textwidth}
<<results_box_crpss_gaussian, fig=TRUE, echo=FALSE, height=3, width=7>>=
# crps skill score by distribution (reference: gaussian)
means$crps_ss_dist <- means$crps
means$crps_ss_dist[means$distribution == "gaussian"] <- 
  1 - means$crps[means$distribution == "gaussian"] / means$crps[means$distribution == "gaussian"]
means$crps_ss_dist[means$distribution == "logistic"] <- 
  1 - means$crps[means$distribution == "logistic"] / means$crps[means$distribution == "gaussian"]
means$crps_ss_dist[means$distribution == "hgaussian"] <- 
  1 - means$crps[means$distribution == "hgaussian"] / means$crps[means$distribution == "gaussian"]

means_sel <- means[means$distribution %in% c("logistic", "hgaussian"),]
means_sel$distribution <- factor(means_sel$distribution, 
                                 levels(means_sel$distribution)[c(2:3)])

# CRPS skill score for distributional forest
s_df <- matrix(ncol = 2, nrow = 15)
s_df[,1] <- means_sel[means_sel$method == "Distributional forest" & means_sel$distribution == "hgaussian", "crps_ss_dist"]
s_df[,2] <- means_sel[means_sel$method == "Distributional forest" & means_sel$distribution == "logistic", "crps_ss_dist"]

# CRPS skill score for prespecified GAMLSS
s_g <- matrix(ncol = 2, nrow = 15)
s_g[,1] <- means_sel[means_sel$method == "Prespecified GAMLSS" & means_sel$distribution == "hgaussian", "crps_ss_dist"]
s_g[,2] <- means_sel[means_sel$method == "Prespecified GAMLSS" & means_sel$distribution == "logistic", "crps_ss_dist"]

# CRPS skill score for boosted GAMLSS
s_gb <- matrix(ncol = 2, nrow = 15)
s_gb[,1] <- means_sel[means_sel$method == "Boosted GAMLSS" & means_sel$distribution == "hgaussian", "crps_ss_dist"]
s_gb[,2] <- means_sel[means_sel$method == "Boosted GAMLSS" & means_sel$distribution == "logistic", "crps_ss_dist"]

# CRPS skill score for EMOS
s_emos <- matrix(ncol = 2, nrow = 15)
s_emos[,1] <- means_sel[means_sel$method == "EMOS" & means_sel$distribution == "hgaussian", "crps_ss_dist"]
s_emos[,2] <- means_sel[means_sel$method == "EMOS" & means_sel$distribution == "logistic", "crps_ss_dist"]

colnames(s_df) <- colnames(s_g) <- colnames(s_gb) <- colnames(s_emos) <- c("hgaussian", "logistic")

# boxplots with matplot
par(mfrow = c(1,4), mar = c(4,4,3,0.7))
matplot(t(s_df[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE,  main = "Distributional forest",
        xlab = "", ylab = "CRPS skill score", xlim = c(0.5, 2.5),
        ylim = c(-0.10, 0.05))
boxplot(s_df, add = TRUE, col = "transparent")
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "Distributional forest"), main = "Distributional forest",
#        ylim = c(-0.12,0.07), las = 2, ylab = "CRPS skill score")
#abline(h = 0, col = pal[5], lwd = 2)
par(mar = c(4,2.9,3,1.8))
matplot(t(s_g[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE, main = "Prespecified GAMLSS",
        xlab = "", ylab = "", xlim = c(0.5, 2.5),
        ylim = c(-0.10, 0.05))
boxplot(s_g, add = TRUE, col = "transparent", yaxt = 'n')
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "Prespecified GAMLSS"), main = "Prespecified GAMLSS",
#        ylim = c(-0.12,0.07), las = 2)
#abline(h = 0, col = pal[5], lwd = 2)
par(mar = c(4,1.8,3,2.9))
matplot(t(s_gb[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE, main = "Boosted GAMLSS",
        xlab = "", ylab = "", xlim = c(0.5, 2.5),
        ylim = c(-0.10, 0.05))
boxplot(s_gb, add = TRUE, col = "transparent", yaxt = 'n')
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "Boosted GAMLSS"), main = "Boosted GAMLSS",
#        ylim = c(-0.12,0.07), las = 2)
#abline(h = 0, col = pal[5], lwd = 2)
par(mar = c(4,0.7,3,4))
matplot(t(s_emos[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE,  main = "EMOS",
        xlab = "", ylab = "", xlim = c(0.5, 2.5),
        ylim = c(-0.10, 0.05))
boxplot(s_emos, add = TRUE, col = "transparent", yaxt = 'n')
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "EMOS"), main = "EMOS",
#        ylim = c(-0.12,0.07), las = 2)
#abline(h = 0, col = pal[5], lwd = 2)
@
\caption{\label{fig:box_crpss_gaussian} CRPS skill score values of each of 
the 15 observation stations, separated by method and distribution. Left-censored 
at zero Gaussian is chosen as reference distribution.}
\end{figure}

@@

\begin{figure}
\setkeys{Gin}{width=0.99\textwidth}
<<results_box_crpss_EMOS, fig=TRUE, echo=FALSE, height=3, width=7>>=
# crps skill score by method (reference: EMOS)
means$crps_ss_method <- means$crps
means$crps_ss_method[means$method == "Distributional forest"] <- 
  1 - means$crps[means$method == "Distributional forest"] / means$crps[means$method == "EMOS"]
means$crps_ss_method[means$method == "Prespecified GAMLSS"] <- 
  1 - means$crps[means$method == "Prespecified GAMLSS"] / means$crps[means$method == "EMOS"]
means$crps_ss_method[means$method == "Boosted GAMLSS"] <- 
  1 - means$crps[means$method == "Boosted GAMLSS"] / means$crps[means$method == "EMOS"]
means$crps_ss_method[means$method == "EMOS"] <- 
  1 - means$crps[means$method == "EMOS"] / means$crps[means$method == "EMOS"]

@@

# CRPS skill score for gaussian
s_gaussian <- matrix(ncol = 3, nrow = 15)
s_gaussian[,1] <- means_sel[means_sel$method == "Distributional forest" & means_sel$distribution == "gaussian", "crps_ss_method"]
s_gaussian[,2] <- means_sel[means_sel$method == "Prespecified GAMLSS" & means_sel$distribution == "gaussian", "crps_ss_method"]
s_gaussian[,3] <- means_sel[means_sel$method == "Boosted GAMLSS" & means_sel$distribution == "gaussian", "crps_ss_method"]

@@

# CRPS skill score for prespecified logistic
s_logistic <- matrix(ncol = 3, nrow = 15)
s_logistic[,1] <- means_sel[means_sel$method == "Distributional forest" & means_sel$distribution == "logistic", "crps_ss_method"]
s_logistic[,2] <- means_sel[means_sel$method == "Prespecified GAMLSS" & means_sel$distribution == "logistic", "crps_ss_method"]
s_logistic[,3] <- means_sel[means_sel$method == "Boosted GAMLSS" & means_sel$distribution == "logistic", "crps_ss_method"]

@@

par(mfrow = c(1,3), mar = c(5,4,3,0))
matplot(t(s_gaussian[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE,  main = "gaussian", ylim = c(-0.12,0.23), 
        xlab = "", ylab = "CRPS skill score", xlim = c(0.5, 3.5))
boxplot(s_gaussian, add = TRUE, col = "transparent", axes = FALSE)
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_method~method, data = means_sel, 
#        subset = (distribution == "gaussian"), main = "gaussian",
#        ylim = c(-0.12,0.23), axes = FALSE, ylab = "CRPS skill score")
axis(1, 0:4, c("",
               "Distr.
                 forest",
               "Presp.
                 GAMLSS",
               "Boosted
                 GAMLSS", ""), las=2)
axis(2, seq(-0.15, 0.25, 0.05), c(seq(-0.15, 0, 0.05), seq(0.05, 0.25, 0.05)), las = 0)
axis(3, 0:4, lwd.ticks = 0, labels = FALSE)
axis(4, seq(-0.15, 0.25, 0.05), lwd.ticks = 0, labels = FALSE)

@@

\begin{figure}
\setkeys{Gin}{width=0.99\textwidth}
<<results_box_crps, fig=TRUE, echo=FALSE, height=3, width=7>>=

# CRPS for gaussian
c_gaussian <- matrix(ncol = 4, nrow = 15)
c_gaussian[,1] <- means[means$method == "Distributional forest" & means$distribution == "gaussian", "crps"]
c_gaussian[,2] <- means[means$method == "Prespecified GAMLSS" & means$distribution == "gaussian", "crps"]
c_gaussian[,3] <- means[means$method == "Boosted GAMLSS" & means$distribution == "gaussian", "crps"]
c_gaussian[,4] <- means[means$method == "EMOS" & means$distribution == "gaussian", "crps"]

>>>>

\item \emph{Hurdle model \citep[or two-part model,][]{Cragg:1971,Mullahy:1986}:}
An additional parameter~$\nu$ describing the probability that there is 
any precipitation at all, i.e., $\nu = \mathbb{P}(Y>0)$, is included in the model.
The three-parameter log-likelihood function for the full hurdle model 
with parameter vector $\bm{\theta} = (\mu, \sigma, \nu)$ is
\begin{equation} \label{eq:hurdle}
\ell_{\mathrm{hurdle}}(\mu, \sigma, \nu; Y) = 
\begin{cases}
\log\left\{\nu \cdot \frac{f(Y; \mu, \sigma)}{1 - F(0; \mu, \sigma)} \right\}, & \text{if } Y > 0 \\[0.2cm]
\log(1 - \nu), & \text{if } Y = 0.
\end{cases}
\end{equation}
For $Y=0$ the probability for no rain is $1-\nu$. 
For $Y>0$ the probability $\nu$ for no rain is multiplied with the 
PDF, left-truncated at zero. In contrast to the censored model, 
there is thus a dedicated parameter $\nu$ controlling the point mass while 
the other parameters $\mu$ and $\sigma$ only influence the distribution of
the positive observations.

\end{itemize}
%
Note that, due to the logs in the first part of Equation~\ref{eq:hurdle},
the log-likelihood of the hurdle model can be separated additively into
a part that depends only on $\nu$ (for all observations) and a part that
depends on $\mu$ and $\sigma$ (for only the positive observations). Therefore,
in a GAMLSS approach, the log-likelihood of the $\nu$ submodel can be
maximized separately from the log-likelihood of the $\mu$ and $\sigma$
submodels to maximize the overall log-likelihood. Hence in all GAMLSS
models considered subsequently a binary logit model is employed for
$Y = 0$ vs.\ $Y > 0$ and a zero-truncated Gaussian model in the case
of $Y > 0$. Similarly, for the distributional forest two separate forests
are employed with a binary and a zero-truncated Gaussian response,
respectively. For comparison with the zero-censored distributions
the combined three-parameter hurdle model as formulated in Equation~\ref{eq:hurdle}
is employed.

All models are evaluated using the continuous ranked probability 
score (CRPS). For the censored models the closed form formulas provided by
\cite{Jordan:2018} are used while a numerical approximation is 
employed for the hurdle models.

\section{Stations}

Out of the 95~observation stations 15 are considered for this 
supplemental study (see Figure~\ref{fig:map}). They have been selected to cover a wide range regarding
geographical location, altitude, and which of the competing models performed
best in the single-split setting reported in Section~3.4 of the main manuscript.

Selected stations: Axams, Lech, Zuers, See im Paznaun, Jungholz, Ladis-Neuegg, 
Oetz, Ochsengarten-Obergut, Ginzling, Rotholz, Walchsee, Koessen, 
Innervillgraten, Matrei in Osttirol, St.~Johann im Walde. See also Supplement~2
(``Stationwise Evaluation'') for more details.

\begin{figure}[t!]
\vspace*{-3.8cm}
\setkeys{Gin}{width=0.99\textwidth}
<<map_stationwise, fig=TRUE, echo=FALSE>>=
##  map
data("StationsTyrol", package = "RainTyrol")
data("MapTyrol", package = "RainTyrol")
library("sp")
sp <- SpatialPointsDataFrame(subset(StationsTyrol, select = c(lon, lat)),
                             data = subset(StationsTyrol, select = -c(lon, lat)),
                             proj4string = raster::crs(MapTyrol$RasterLayer))

plot(MapTyrol$SpatialPolygons)
points(sp, pch = 21, col = "darkgrey", las = 1, cex = 0.6)
points(sp[c(5, 6, 20, 23, 25, 32, 33, 46, 47, 56, 57, 70, 79, 82, 83),], 
       pch = 21, bg = hcl(325, 100, 70), cex = 1.5)
# stationname beneath
text(sp[c(25, 47, 57, 70),], 
     labels = StationsTyrol$name[c(25, 47, 57, 70)],
     pos=1, cex=0.75)
# stationname left
text(sp[c(20, 32, 46, 79, 83),], 
     labels = StationsTyrol$name[c(20, 32, 46, 79, 83)],
     pos=2, cex=0.75)
# stationname above
text(sp[c(6, 33, 82),], 
     labels = StationsTyrol$name[c(6, 33, 82)],
     pos=3, cex=0.75)
# stationname right
text(sp[c(5, 23, 56),], 
     labels = StationsTyrol$name[c(5, 23, 56)],
     pos=4, cex=0.75)
@
\vspace*{-3.8cm}
\caption{\label{fig:map}Map of Tyrol with all 95 observation stations and 
the 15 considered stations highlighted in pink.}
\end{figure}

\pagebreak

\section{Results}

\subsubsection*{CRPS for all distributions by model and station}

Each panel of Figure~\ref{fig:lattice} compares the CRPS across the three distributions
considered: \textit{cgaussian} (zero-censored Gaussian), \textit{hgaussian} (Gaussian hurdle model),
and \textit{clogistic} (zero-censored logistic). The four different models considered
are coded by both colors and symbols and connected by lines for easier comparison between
models. As the average CRPS (i.e., the difficulty of the prediction problem) varies
substantially by station, separate panels are employed for each station.
%
\begin{itemize}

\item For the distributional forest all three distributions perform very similarly for all
stations. Thus, it appears to be very robust against missspecifications of the distribution.

\item Similarly, for the EMOS model all distributions lead to comparable CRPS for all
stations. However, there is some more variation in CRPS across distributions compared to
the distributional forest.

\item In contrast, the two GAMLSS models typically perform better (i.e., have lower CRPS) for
the Gaussian distributions compared to the logistic distribution. The preference of
censored vs.\ hurdle specification is not unanimous across stations.

\end{itemize}
%
\textit{Note:} When applying the hurdle model at stations Oetz and 
Innvervillgraten the prespecified GAMLSS could not be estimated (using the \proglang{R}
package \pkg{gamlss}) due to numerical issues in the fitting algorithm.
Therefore, this method is not represented in the two corresponding panels 
of Figure~\ref{fig:lattice}.

\begin{figure}[p!]
\setkeys{Gin}{width=0.99\textwidth}
<<results_lattice, fig=TRUE, echo=FALSE, results=hide, height=10, width=8>>=
library("lattice")
##### 
# HCL palette
pal <- hcl(c(10, 128, 260, 290, 50), 100, 50)

load("Rain_distributions.rda")
levels(results$means$distribution) <- c("cgaussian", "hgaussian", "clogistic")
levels(results$means$station)[16] <- "St. Johann im Walde"

stations <- c("Axams", "Lech", "Zuers", "See im Paznaun", "Jungholz", 
              "Ladis-Neuegg", "Oetz", "Ochsengarten-Obergut",
              "Ginzling", "Rotholz", "Walchsee", "Koessen", 
              "Innervillgraten", "Matrei in Osttirol", 
              "St. Johann im Walde")

@@

xyplot(crps ~ distribution | station, groups = ~ method, 
       data = means, auto.key = TRUE, 
       type = "o", lwd = 2, lty = 1,  
       drop.unused.levels = TRUE, layout = c(3,5),
       as.table = TRUE, ylab = "CRPS", xlab = "Distribution")
@
\caption{\label{fig:lattice}CRPS values (lower = better) across the three
distributions for all four models and 15~stations.}
\end{figure}

\subsubsection*{CRPS skill score in reference to zero-censored Gaussian}

To aggregate the CRPS results across stations, CRPS skill scores are visualized
with boxplots in Figure~\ref{fig:box_crpss_gaussian}, using the zero-censored Gaussian distribution
as the reference distribution (orange horizontal line at zero). The underlying skill
scores for each station are also included in the graphic by gray lines.
%
\begin{itemize}

\item As before, there are only slight variations for the distributional forest, mostly
  between $\pm 1$\%. While the median CRPS skill score over the 15 stations for the 
  Gaussian hurdle model is slightly better than for the simpler zero-censored Gaussian
  model, this advantage is very small and not systematic across stations.

\item For EMOS, the median skill score for both Gaussian hurdle and zero-censored logistic
  are very close to the zero-censored Gaussian reference. However, the variation for the
  hurdle model is much higher than for the censored model.

\item For the two GAMLSS models the Gaussian hurdle is quite close to the zero-censored
  Gaussian reference, while the zero-censored logistic clearly performs worse (especially
  in the prespecified GAMLSS).

\begin{figure}[t!]
\setkeys{Gin}{width=0.99\textwidth}
<<results_box_crpss_gaussian, fig=TRUE, echo=FALSE, height=3, width=7.2>>=
# crps skill score by distribution (reference: gaussian)
means$crps_ss_dist <- means$crps
means$crps_ss_dist[means$distribution == "cgaussian"] <- 
  1 - means$crps[means$distribution == "cgaussian"] / means$crps[means$distribution == "cgaussian"]
means$crps_ss_dist[means$distribution == "clogistic"] <- 
  1 - means$crps[means$distribution == "clogistic"] / means$crps[means$distribution == "cgaussian"]
means$crps_ss_dist[means$distribution == "hgaussian"] <- 
  1 - means$crps[means$distribution == "hgaussian"] / means$crps[means$distribution == "cgaussian"]

means_sel <- means[means$distribution %in% c("clogistic", "hgaussian"),]
means_sel$distribution <- factor(means_sel$distribution, 
                                 levels(means_sel$distribution)[c(2:3)])

# CRPS skill score for distributional forest
s_df <- matrix(ncol = 2, nrow = 15)
s_df[,1] <- means_sel[means_sel$method == "Distributional forest" & means_sel$distribution == "hgaussian", "crps_ss_dist"]
s_df[,2] <- means_sel[means_sel$method == "Distributional forest" & means_sel$distribution == "clogistic", "crps_ss_dist"]

# CRPS skill score for prespecified GAMLSS
s_g <- matrix(ncol = 2, nrow = 15)
s_g[,1] <- means_sel[means_sel$method == "Prespecified GAMLSS" & means_sel$distribution == "hgaussian", "crps_ss_dist"]
s_g[,2] <- means_sel[means_sel$method == "Prespecified GAMLSS" & means_sel$distribution == "clogistic", "crps_ss_dist"]

# CRPS skill score for boosted GAMLSS
s_gb <- matrix(ncol = 2, nrow = 15)
s_gb[,1] <- means_sel[means_sel$method == "Boosted GAMLSS" & means_sel$distribution == "hgaussian", "crps_ss_dist"]
s_gb[,2] <- means_sel[means_sel$method == "Boosted GAMLSS" & means_sel$distribution == "clogistic", "crps_ss_dist"]

# CRPS skill score for EMOS
s_emos <- matrix(ncol = 2, nrow = 15)
s_emos[,1] <- means_sel[means_sel$method == "EMOS" & means_sel$distribution == "hgaussian", "crps_ss_dist"]
s_emos[,2] <- means_sel[means_sel$method == "EMOS" & means_sel$distribution == "clogistic", "crps_ss_dist"]

colnames(s_df) <- colnames(s_g) <- colnames(s_gb) <- colnames(s_emos) <- c("hgaussian", "clogistic")

# boxplots with matplot
par(mfrow = c(1,4), mar = c(4,4,3,0.7))
matplot(t(s_df[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE,  main = "Distributional forest",
        xlab = "", ylab = "CRPS skill score", xlim = c(0.5, 2.5),
        ylim = c(-0.10, 0.05))
boxplot(s_df, add = TRUE, col = "transparent")
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "Distributional forest"), main = "Distributional forest",
#        ylim = c(-0.12,0.07), las = 2, ylab = "CRPS skill score")
#abline(h = 0, col = pal[5], lwd = 2)
par(mar = c(4,2.9,3,1.8))
matplot(t(s_g[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE, main = "Prespecified GAMLSS",
        xlab = "", ylab = "", xlim = c(0.5, 2.5),
        ylim = c(-0.10, 0.05))
boxplot(s_g, add = TRUE, col = "transparent", yaxt = 'n')
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "Prespecified GAMLSS"), main = "Prespecified GAMLSS",
#        ylim = c(-0.12,0.07), las = 2)
#abline(h = 0, col = pal[5], lwd = 2)
par(mar = c(4,1.8,3,2.9))
matplot(t(s_gb[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE, main = "Boosted GAMLSS",
        xlab = "", ylab = "", xlim = c(0.5, 2.5),
        ylim = c(-0.10, 0.05))
boxplot(s_gb, add = TRUE, col = "transparent", yaxt = 'n')
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "Boosted GAMLSS"), main = "Boosted GAMLSS",
#        ylim = c(-0.12,0.07), las = 2)
#abline(h = 0, col = pal[5], lwd = 2)
par(mar = c(4,0.7,3,4))
matplot(t(s_emos[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE,  main = "EMOS",
        xlab = "", ylab = "", xlim = c(0.5, 2.5),
        ylim = c(-0.10, 0.05))
boxplot(s_emos, add = TRUE, col = "transparent", yaxt = 'n')
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "EMOS"), main = "EMOS",
#        ylim = c(-0.12,0.07), las = 2)
#abline(h = 0, col = pal[5], lwd = 2)
@
\caption{\label{fig:box_crpss_gaussian} CRPS skill score (higher = better)
for hurdle Gaussian and censored logistic distributions in reference to the
censored Gaussian distribution. Lines correspond to the 15 stations and are
aggregated in the boxplot. The y-axis range is limited
to $-10$\%, lower values for the two GAMLSS models are not displayed.}
\end{figure}

@@

\begin{figure}
\setkeys{Gin}{width=0.99\textwidth}
<<results_box_crpss_EMOS, fig=TRUE, echo=FALSE, height=3, width=7.2>>=
# crps skill score by method (reference: EMOS)
means$crps_ss_method <- means$crps
means$crps_ss_method[means$method == "Distributional forest"] <- 
  1 - means$crps[means$method == "Distributional forest"] / means$crps[means$method == "EMOS"]
means$crps_ss_method[means$method == "Prespecified GAMLSS"] <- 
  1 - means$crps[means$method == "Prespecified GAMLSS"] / means$crps[means$method == "EMOS"]
means$crps_ss_method[means$method == "Boosted GAMLSS"] <- 
  1 - means$crps[means$method == "Boosted GAMLSS"] / means$crps[means$method == "EMOS"]
means$crps_ss_method[means$method == "EMOS"] <- 
  1 - means$crps[means$method == "EMOS"] / means$crps[means$method == "EMOS"]

@@

# CRPS skill score for gaussian
s_gaussian <- matrix(ncol = 3, nrow = 15)
s_gaussian[,1] <- means_sel[means_sel$method == "Distributional forest" & means_sel$distribution == "cgaussian", "crps_ss_method"]
s_gaussian[,2] <- means_sel[means_sel$method == "Prespecified GAMLSS" & means_sel$distribution == "cgaussian", "crps_ss_method"]
s_gaussian[,3] <- means_sel[means_sel$method == "Boosted GAMLSS" & means_sel$distribution == "cgaussian", "crps_ss_method"]

@@

# CRPS skill score for prespecified logistic
s_logistic <- matrix(ncol = 3, nrow = 15)
s_logistic[,1] <- means_sel[means_sel$method == "Distributional forest" & means_sel$distribution == "clogistic", "crps_ss_method"]
s_logistic[,2] <- means_sel[means_sel$method == "Prespecified GAMLSS" & means_sel$distribution == "clogistic", "crps_ss_method"]
s_logistic[,3] <- means_sel[means_sel$method == "Boosted GAMLSS" & means_sel$distribution == "clogistic", "crps_ss_method"]

@@

par(mfrow = c(1,3), mar = c(5,4,3,0))
matplot(t(s_gaussian[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE,  main = "cgaussian", ylim = c(-0.12,0.23), 
        xlab = "", ylab = "CRPS skill score", xlim = c(0.5, 3.5))
boxplot(s_gaussian, add = TRUE, col = "transparent", axes = FALSE)
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_method~method, data = means_sel, 
#        subset = (distribution == "cgaussian"), main = "cgaussian",
#        ylim = c(-0.12,0.23), axes = FALSE, ylab = "CRPS skill score")
axis(1, 0:4, c("",
               "Distr.
                 forest",
               "Presp.
                 GAMLSS",
               "Boosted
                 GAMLSS", ""), las=2)
axis(2, seq(-0.15, 0.25, 0.05), c(seq(-0.15, 0, 0.05), seq(0.05, 0.25, 0.05)), las = 0)
axis(3, 0:4, lwd.ticks = 0, labels = FALSE)
axis(4, seq(-0.15, 0.25, 0.05), lwd.ticks = 0, labels = FALSE)

@@

\begin{figure}
\setkeys{Gin}{width=0.99\textwidth}
<<results_box_crps, fig=TRUE, echo=FALSE, height=3, width=7.2>>=

# CRPS for gaussian
c_gaussian <- matrix(ncol = 4, nrow = 15)
c_gaussian[,1] <- means[means$method == "Distributional forest" & means$distribution == "cgaussian", "crps"]
c_gaussian[,2] <- means[means$method == "Prespecified GAMLSS" & means$distribution == "cgaussian", "crps"]
c_gaussian[,3] <- means[means$method == "Boosted GAMLSS" & means$distribution == "cgaussian", "crps"]
c_gaussian[,4] <- means[means$method == "EMOS" & means$distribution == "cgaussian", "crps"]
